\section{Dynamic Class Loading}
Dynamic class loading is used in several places in sandmark to allow
for extensibility.  The major problem in dynamic class loading in sandmark
is finding classes to load.  In particular, sandmark can be run from a 
directory or a jar file, and it may or may not know the location of that 
directory or jar file or even which of the two it is running from.  

All of this detail is abstracted by sandmark.util.classloading.ClassFinder.
ClassFinder allows a caller to specify one of a small number of classes 
(defined in sandmark.util.classloading.IClassFinder), and
ClassFinder will return the names of all classes derived from that class.  For example,
the following code prints to standard output all classes derived from
sandmark.Algorithm.

\begin{listing}{1}
java.util.Collection c = 
   sandmark.util.classloading.ClassFinder.getClassesWithAncestor
      (sandmark.util.classloading.IClassFinder.ALGORITHM);
java.util.Iterator it = c.iterator();
while(it.hasNext()) {
   String className = (String)c.next();
   System.out.println(className);
}
\end{listing}

The class names returned by ClassFinder are those returned by
an implementor of \\sandmark.util.classloading.IClassFinder.  ClassFinder 
uses the first implementation in the following list
whose default constructor does not throw an exception:

\begin{enumerate}
\item \textbf{FileClassFinder} This class reads class names and from which 
of the interfaces in \\sandmark.util.classloading.IClassFinder they derive
from a file call \textbf{Algorithms.txt}.  This file is opened by a call to \\
\textbf{ClassLoader.getSystemClassLoader().getResourceAsStream(``Algorithms.txt'')}.
If this call fails, FileClassFinder's default constructor throws an exception.

\item \textbf{JarClassFinder} This class searches through all entries in the
jar file named by \\System.getProperty(``SMARK\_PATH''), 
loads all class files contained therein, and determines from 
which classes in IClassFinder each of these classes derives. If opening
the jar file for reading throws an exception, JarClassFinder's default 
constructor throws an exception.

\item \textbf{DirClassFinder} This class is similar to JarClassFinder, but 
it searches through the file system hierarchy rooted at 
System.getProperty(``SMARK\_ROOT'').  Its default constructor throws an exception
if the contents of SMARK\_ROOT cannot be read.
\end{enumerate}








